\documentclass{book}
\usepackage[utf8]{inputenc}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools}


\begin{document}

\chapter{\emph{Probability}}
%\chapter{\emph{Probability}} 

\label{probability}

\section{What is probability?}
Kolmogorov axioms

\section{What is it, really?}

\section{Random variables}

\section{Independence and Bayes theorem}
\subsection{Illustrating Bayes Theorem}
%In this section I am assuming that someone will transcript the example seen in class as an example in the book 

At first sight, the Bayes Theorem don't appear so intuitive. But we can illustrate it using the set theorem.

Firstly let's rescue the last example. A test 99\% precise to a sickness prevalent in 0.01\% of the population. Let A be a set of 1,000,000 persons. the subset of persons with this sickness B must have 100 elements.

Suppose now that we applied the test to all persons in set A. We can divide this set in 4 subsets:

$TP \subset A$: \{the test results a positive and they are sick\}

$FP \subset A$:\{the test results a positive, but they aren't sick\}

$TN \subset A$:\{the test results a negative and they aren't sick\}

$FN \subset A$:\{the test results a negative, but they are sick\}

So if we have 999,900 folk healthy than $99\%*999,900 = 989,901$ will result negative and 9,999 will result positive. And from our 100 folks sick 99 will result positive in test and 1 negative.

So our sets will be:

$||TN|| = 989,901$

$||FN|| = 1$

$||TP|| = 99$

$||FP|| = 9,999$

When we as what is the odds of receiving a positive and being sick we are looking to the subset of people who receive positive in the test that is $FP \cap TP$. And thus, considering all elements equiprobable $\Pr$(receiving positive and being sick)=$\frac{TP}{FP \cap TP} = \frac{99}{99+9,999}=0.0098$. Exacly the result founded by the Bayes Theorem.


\subsection{Solved Exercises}

%\title{ExercisePregnancy}
%\author{Carmen Melo Toledo}


\subsubsection{Pregnancy Test}

    According to CDC (Center for Disease Control and Prevetion):
    

    \begin{center}
     \begin{tabular}{||c | c||} 
     \hline
     Contraceptive Method & Percentage of failure\footnote{percentage of undesired pregnancy after a year} \\ [0.9ex] 
     \hline\hline
     Male Codon & 18\% \\ 
    \hline
    Pill & 9\% \\ 
    \hline
    Fertility-Awareness based method & 24\% \\ 
    \hline
    IUD & 0.5\% \\
    \hline
    \end{tabular}
    \end{center}
    consider a pregnancy test that 99\% accurate to positive and to negative\footnote{pregnancy tests usually are less accurate to avoid fake negatives, but to exemplify let's assume it}  and estimated the percentage of having a typical fertile couple have a fake positive after a year using:
    \begin{enumerate}[(a)]
        \item Male Codon
        \item Pill
        \item IUD
        \item Pill and Male Codon
    \end{enumerate}
    


\textbf{Answer}
To solve this problem we shall use tow equations that we saw earlier
\large{Bayes: }
\begin{equation*}
 \Pr[a|b]=\frac{\Pr[b|a]*\Pr[a]}{\Pr[b]}
\end{equation*}


 \large{theorem of total probability:}
\begin{equation*}
    \\ \Pr[a]=\Pr[a|b]*Pr[b]+Pr[a|\sim b]*Pr[\sim b]
\end{equation*}

So in general we have:

\begin{equation*}
    \large{\Pr[not pregnant| positive]=\frac{\Pr[positive|not pregnant]*\Pr[not pregnant]}{\Pr[positive]}}
\end{equation*}

where $\Pr[positive|not pregnant]$  is  the  accuracy,

$\Pr[not pregnant]$ is 100\% - chance of failure and 

$\Pr[positive]$ will be given by the second theorem

\begin{equation*}
\large{\Pr[positive]=\Pr[Positive|not pregnant}*\Pr[not pregnant]+\Pr[Positive|pregnant]*\Pr[pregnant]    
\end{equation*}

Let's apply to each letter:

a)

$\Pr[positive|not pregnant] = 1\%$

$\Pr[not pregnant]= (100-18)\%=82\%$

$\Pr[positive]=0.01*0.82+0.99*0.18=0.1864$

So we have

$\Pr[fake positive] = \frac{0.01*0.82}{0.1864}=0.044 or 4.4\%$


b)

$\Pr[positive|not pregnant] = 1\%$

$\Pr[not pregnant]= (100-9)\%=91\%$

$\Pr[positive]=0.01*0.91+0.99*0.09=0,0982$

So we have

$\Pr[fake positive] = \frac{0.01*0.91}{0.0982}=0.09266 or 9.3\%$

c)

$\Pr[positive|not pregnant] = 0.5\%$

$\Pr[not pregnant]= (100-0.5)\%=99.5\%$

$\Pr[positive]=0.01*0.995+0.99*0.005=0,0149$

So we have

$\Pr[fake positive] = \frac{0.01*0.995}{0.0149}=0.6678=66.8\%$

d)

Using codon and taking pill are independent events so the failure rate is the multiplication of the failure rates

Failure Rate: 0.0162 = 1.62\%

$\Pr[positive|not pregnant] = 1.62\%$

$\Pr[not pregnant]= (100-1.62)\%=98.38\%$

$\Pr[positive]=0.01*0.9838+0.99*0.0162=0,02588$

So we have

$\Pr[fake positive] = \frac{0.01*0.9838}{0.02588}=0.3801=38\%$

\subsubsection{exercise 2}


\section{Probability models}


\end{document}